\section{Implementation}\label{sec:implementation}
In this section we describe how we implemented the presented syntactical and semantical system
in Why3. This is presented for each syntactical component: arithmetic expressions, boolean expression, statements, and formulas. In \ref{sec:evaluator} we describe how we implemented
an evaluator for the language, and in \ref{sec:vcg} the implementation of verification
condition generation is presented.

For each component we define a set of properties which must hold for a correct implementation.
We sometimes accompany the Why3 implementations with proofs conducted in Isabelle, to reason
about parts that were tricky in Why3.

The entire implemetation can be seen in \cref{code}.

\subsection{Syntax}
\input{Implementation/syntax}

\subsection{Arithmetic Expressions}
In the following we will describe how the semantics for arithmetic expressions are implemented,
and which properties we want to hold to ensure correctness of the implementation.

\subsubsection{Semantics}
\input{Implementation/asemantics}

\subsubsection{Properties}\label{sec:aprops}
\input{Implementation/aproperties}

\subsection{Boolean Expressions}
Now we will describe the implementation of semantics for boolean expressions, together with
the properties we want to ensure holds.

\input{Implementation/bsemantics}

\subsection{Statements}
Finally we present our implementation of semantics of statements.
\input{Implementation/ssemantics}


\subsection{Assertion Language (Formulas)}\label{sec:iformulas}
\input{Implementation/formulas}

\subsection{Evaluator}\label{sec:evaluator}
\input{Implementation/evaluator}

\subsubsection{Modelling a store}\label{sec:model}
\input{Implementation/model}

\subsubsection{Evaluation}
\input{Implementation/evaluation}

\subsection{Verification Condition Generation}\label{sec:vcg}
\input{Implementation/vcgen}

\subsubsection{Variable substitution}
\input{Implementation/subst}

\subsubsection{Weakest Liberal Precondtion}\label{sec:impwlp}
\input{Implementation/wlp}

\section{Extraction of code}\label{sec:extract}
\input{Implementation/extract}
