From \autoref{TODO} it should be clear that there are multiple situations where we want to substitute
all free occurences of a variable by an new variable.
For instance when $s$ is $x := a$ in $wlp(s, Q)$ where $Q$ is a formula, and $s$ is the assignment of $a$ to $x$,
we want to update each occurence of $x$ in $Q$, but do not substitute bound instances of $x$.
The rules are as follows, we only consider cases where variables might occur. $f_{i}$ is formulas, $x,y,z$ are variables and $b_{i}, a_{i}$ describes boolean expression and arithmetic expressions respectively. Conjuction for booleans and formulas are equivalent.

\begin{align*}
\subst{x}{x}{y} &= y \\
\subst{z}{x}{y} &= x \\
&\\
\subst{z}{x}{y} &= x \\
\subst{(a_{1} \leq a_{2})}{x}{y} &= \subst{a_{1}}{x}{y} \le \subst{a_{2}}{x}{y} \\
\subst{(b_{1} \wedge b_{2})}{x}{y} &= \subst{b_{1}}{x}{y} \wedge \subst{b_{2}}{x}{y} \\
\subst{(\neg b)}{x}{y} &= \neg\subst{b}{x}{y} \\
&\\
\subst{(f_{1} \Rightarrow f_{2})}{x}{y} &= \subst{f_{1}}{x}{y} \Rightarrow \subst{f_{2}}{x}{y} \\
\subst{(\forall x. f)}{x}{y} &=  \forall x. f\\
\subst{(\forall z. f)}{x}{y} &=  \forall z. \subst{f}{x}{y}\\
\end{align*}
The $subst$ must also adhere to the rule that $y$ must be free for $x$ in whatever the expression/formula,
meaning a free occurence of $x$ must not be bound when substituted by $y$.

From the formulation of the substitution function it should be easy to see, that we recurse down the syntax tree.
The most interesting case is when we meet a quantifier. If the variable $x$ we want to substitute is bound then we stop recursion, as all occurences of $x$ will then be bound.

It should be noted that we do not ensure that $x$ must be free for $y$,
thus the function is rather unsafe.
We only ever use it in a context where this cannot happen, since we instantiate a new variable that does not occur in the formula, bound or free. The variable is so-called fresh in the formula.

The way we generate a fresh variable, is by traversing the syntax tree of the formula and recursively taking the maximal variable value and adding 1.
This works since we use integers as identifiers.
A similar result can be obtained by all infinite enumerable sets. For instance if identifiers were strings, one could add a new character to the longest variable name.

This seems rather trivial, however we formulated predicates that state whether a variable is fresh in an expression or formula. We then afterward tried making a lemmas which stated:

\begin{lstlisting}
lemma fresh_var_is_fresh : forall f v.
    v = fresh_from f -> fresh_in_formula v f
\end{lstlisting}

The lemma simply asserts that generating a fresh variable $v$ from $f$ implies that $v$ is fresh in $f$.
We got the inspiration for this approach from \cite{TODO: WP revisited}.
In their work they formulate the fresh_from function using axioms.
That is they provide a function declaration and then the function ``computes'' the value based on a set of axioms.
For us this is not a viable approach, as such functions cannot be extracted. Therefore we defined the function ourself and stated axiom as a lemma to ensure that \texttt{fresh\_from} always ensures a fresh variable.
Although this seems like a trivial result, we were not able to prove this using the SMT solvers.
We tried to define this system in isabelle to see how difficult it was to prove using a proof assistant, where one has more control. We succeeded in proving this.
But since, this does not entail that there might be a bug in the why3 code.
The isabelle proof might not be the cleverest way to prove it, but we proved it the following way.
let $V$ be the set of all vars in an arithmetic expression $a$.
Then assuming $v$ is the maximum identifier in $a$, we show that $\forall x \in V. v \ge x$. We do so by induction.
By this we can then show $v + 1 \notin V$.
We then made a lemma stating that given $v \notin V$ then $v$ is fresh in $a$.
From this we can show that $v + 1$ is fresh in $V$.
We then do the same for boolean expression and formulas, and the proof is done.
The full proof can be seen in \autoref{TODO appendix}.